A Bias-Variance Decomposition of Information Transfer in Multi-Task Learning

### 2020-08-17 arxiv/proof_examples.tex
pf

\begin{proof}[Proof of Proposition \ref{prop_data_size}]
Following the proof of Proposition \ref{prop_dist_transition}, we see that $\te(\hat{\beta}_2^{\MTL})< \te(\hat{\beta}_2^{\STL})$ w.h.p. if \eqref{pos1} holds, while $\te(\hat{\beta}_2^{\MTL})>\te(\hat{\beta}_2^{\STL})$ w.h.p. if \eqref{neg1} holds. 
%\be\label{pos1}
%pd^2  \cdot \frac{\rho_1 (\rho_1 + \rho_2) (\rho_2 - 1)}{(\rho_1 + \rho_2 - 1)^2}\bigbrace{1 + \sqrt{\frac{1}{\rho_1}}}^4 <  \sigma^2  \bigbrace{1 - \oo(1)},\ee
%and we get from \eqref{upper102} that $\te(\hat{\beta}_t^{\MTL})>\te(\hat{\beta}_t^{\STL})$ w.h.p. if
%\be\label{neg1}pd^2  \cdot \frac{\rho_1 (\rho_1 + \rho_2) (\rho_2 - 1)}{(\rho_1 + \rho_2 - 1)^2}\bigbrace{1 - \sqrt{\frac{1}{\rho_1}}}^4 >  \sigma^2  \bigbrace{1 + \oo(1)}.\ee
%then we have $\te(\hat{\beta}_t^{\MTL})> \te(\hat{\beta}_t^{\STL})$ w.h.p..

We first explain the meaning of the condition 
\be\label{explainpsi}\Psi(\beta_1,\beta_2)>2/(\rho_2-1).\ee 
%For the first statement of Proposition \ref{prop_data_size}, we 
Notice that the function
$$ \Phi(\rho_1, \rho_2)=\frac{(\rho_1 + \rho_2 - 1)^2}{\rho_1 (\rho_1 + \rho_2) (\rho_2 - 1)}=\frac{1}{\rho_2-1} \left(1 +\frac{\rho_2-2}{\rho_1}+\frac{1}{\rho_1(\rho_1+\rho_2)}\right)$$
is strictly decreasing with respect to $\rho_1$ as long as $\rho_2> 2$, and $ \Phi(\rho_1, \rho_2)$ converges to $(\rho_2-1)^{-1}$ as $\rho_1\to \infty$.  Moreover, we notice that $\left( \al_-^2(\rho_1) -  \oo(1) \right)^{-1} < 2$ for $\rho_1>40$. Hence \eqref{explainpsi} implies that \eqref{neg1} holds for all  large enough $\rho_1$. The transition from positive transfer when $\rho_1$ is small to negative transfer when $\rho_1$ is large is described by the two bounds in Proposition \ref{prop_data_size}.

% $$ \Psi(\beta_1,\beta_2) \ge \bigbrace{1 + \sqrt{\frac{1}{\rho_1}}}^{-4}\frac{\sigma^2}{\rho_2-1}\left(1 -\oo(1)\right),$$ 
% then \eqref{pos1} holds, which shows that  $\te(\hat{\beta}_t^{\MTL}) < \te(\hat{\beta}_t^{\STL})$ holds w.h.p.. Plugging into $\rho_1>50$, we obtain the first statement. ...... Note that for $\rho_1>40$, we have 
%$\gamma_+^{-1}$

The two bounds follows directly from \eqref{pos1} and \eqref{neg1}. We will use the following trivial inequalities 
\be\label{trivialphi}
\frac{(\rho_2-1) \rho_1 }{\rho_1+\rho_2-2}\cdot \left( 1-\frac{1}{(\rho_1+\rho_2-2)^2}\right) \le  \Phi(\rho_1, \rho_2) \le  \frac{(\rho_2-1) \rho_1 }{\rho_1+\rho_2-2}.
\ee

@@

 Note that if we require the RHS of \eqref{addconstraint} to be larger than $40$, that is, \eqref{addconstraint} is not a null condition. Then together with \eqref{explainpsi}, we get
 $$ \rho_2 - 2>\left[ 2\left( \al_+^2(\rho_1) +  \oo(1) \right)  -1\right]\rho_1 .$$
Plugging into $\rho_1>40$, we get $\rho_2 \ge 106$. This gives a constraint on $\rho_2$.

@@

 \noindent{\bf Negative transfer.} 
With \eqref{trivialphi}, we see that \eqref{neg1} is implied by the following inequality:
\begin{align}\label{neg1solv}
 & \Psi(\beta_1,\beta_2)  \frac{(\rho_2-1) \rho_1 }{\rho_1+\rho_2-2}\left( 1-\frac{1}{(\rho_1+\rho_2-2)^2}\right) >  \Psi(\beta_1,\beta_2) \frac{(\rho_2-1.5) \rho_1 }{\rho_1+\rho_2-2}> \left( \al_-^2(\rho_1) -  \oo(1) \right)^{-1} .
 \end{align}
 where we used $(1-(\rho_1+\rho_2-2)^{-2})(\rho_2 - 1) > \rho_2-1.5$ for $\rho_1>40$ and $\rho_2>110$. Then we can solve \eqref{neg1solv} to get
 \be\label{addconstaint2}
 \rho_1 > \frac{(\rho_2-2) \sigma^2}{\Psi(\beta_1, \beta_2) \cdot  (\rho_2 - 1.5)\left( \al_-^2(\rho_1) -  \oo(1) \right)  - 1}  ,
 \ee
which gives the second statement of Proposition \ref{prop_data_size}.
We remark that condition \eqref{explainpsi} implies $\Psi(\beta_1, \beta_2) \cdot  (\rho_2 - 1.5)\left( \al_-^2(\rho_1) -  \oo(1) \right)>1$, so \eqref{addconstaint2} does not give a trivial bound. 
 \end{proof}

Next we prove Proposition \ref{prop_data_efficiency}, which gives upper and lower bounds on the data efficiency ratio for Taskonomy.

\begin{proof}[Proof of Proposition \ref{prop_data_efficiency}]
Suppose we have reduced number of datapoints---$x n_1$ for task 1 and $x n_2$ for task 2 with $n_1=\rho_1 p$ and $n_2=\rho_2 p$. Then all the results in the proof of Proposition \ref{prop_dist_transition} still hold, except that we need to replace $(\rho_1,\rho_2)$ with $(x\rho_1,x\rho_2)$. More precisely, we have
	\begin{align*}
		 a_1 = \frac{\rho_1(x\rho_1 + x\rho_2 - 1)}{x(\rho_1 + \rho_2)^2} ,\quad
		& a_2 = \frac{\rho_2(x\rho_1 + x\rho_2 - 1)}{x(\rho_1 + \rho_2)^2} ,  \\
		 a_3 = \frac{\rho_2}{(\rho_1 + \rho_2)(x\rho_1 + x\rho_2 - 1)}, \quad
		& a_4 = \frac{\rho_1}{(\rho_1 + \rho_2)(x\rho_1 + x\rho_2 - 1)}. 
	\end{align*}
Moreover, with high probability,
	\be\label{reduceproof1}
\begin{split}
\te_i(\hat \beta_i^{\MTL}(x)) =  \frac{\sigma^2}{x (\rho_1+\rho_2) - 1}\left(1+ \oo(1)\right)+ \delta^{(i)}_{\bias}, \quad i=1,2.
\end{split}
\ee
 Here the model shift biases $\delta^{(i)}_{\bias}$ satisfy that  
\be\nonumber %\label{Deltabeta} 
\al_-^2(\al\rho_i) - \oo(1)  \le {\delta_{\bias}^{(i)}}/{ \Delta_{\bias}^{(i)}} \le \al_+^2(\al\rho_i) +  \oo(1) , \quad i=1,2,\ee	
where $ \Delta^{(i)}_{\bias}$ are defined as 
\be \nonumber
\Delta^{(i)}_{\bias} := pd^2 \frac{(x\rho_i)^2\cdot x (\rho_1+\rho_2)}{[x ( \rho_1+\rho_2) - 1]^3} ,\quad i=1,2,. 
\ee
On the other hand, using Lemma \ref{lem_minv} we have w.h.p.,
\be\label{reduceproof2} 
\te_i(\hat{\beta}_i^{\STL}) = \frac{\sigma^2}{\rho_i -1} \left( 1+ \oo(1)\right),\quad i=1,2.
\ee
Comparing \eqref{reduceproof1} and \eqref{reduceproof2}, we immediately obtain the lower bound $x^\star\ge x_l $. 
%where
%$$\al_l:=\frac1{\rho_1+\rho_2}\left(\frac{2}{\frac1{\rho_1-1}+\frac1{\rho_2 -1}}+1\right) \ge \frac{\min(\rho_1,\rho_2)}{\rho_1+\rho_2}.$$
In fact, one can see that if $x< x_l$, then we have 
$$ \frac{2\sigma^2}{x (\rho_1+\rho_2) - 1} > \frac{\sigma^2}{\rho_1-1}+\frac{\sigma^2}{\rho_2-1},$$
that is, $\te_1(\hat{\beta}_1^{\MTL}(x)) + \te_2(\hat{\beta}_2^{\MTL}(x))$ is larger than $\te_1(\hat{\beta}_t^{\STL}) + \te_2(\hat{\beta}_t^{\STL})$ even if we do not take into account the model shift bias terms $ \delta^{(i)}_{\bias}$.

@@

\noindent{\bf The upper bound.} From \eqref{reduceproof1} and \eqref{reduceproof2}, we see that $x^\star\le x$ if $x$ satisfies 
\be\nonumber
\begin{split}
&(1+\oo(1)) \cdot \sum_{i=1}^2 pd^2 \frac{(x\rho_i)^2\cdot x\rho }{(x\rho - 1)^3}  \bigbrace{1 + \sqrt{\frac{1}{x \rho_i}}}^4 \le \frac{\sigma^2}{\rho_1 -1}+\frac{\sigma^2}{\rho_2 -1} - \frac{2\sigma^2}{x\rho  - 1} . 
\end{split}
\ee
We rewrite the inequality as
\be
\begin{split}\label{solval1}
  (1+\oo(1)) \cdot \frac{\Psi(\beta_1,\beta_2)}{[1 - (x\rho)^{-1}]^3} \sum_{i=1}^2 \bigbrace{\sqrt{\frac{\rho_i}{\rho}} + \sqrt{\frac{1}{x\rho}}}^4\le \frac{1}{\rho_1 -1}+\frac{1}{\rho_2 -1} - \frac{2}{x \rho  - 1}.
\end{split}
\ee
With $x\rho\ge \min(\rho_1,\rho_2)\ge 9$, we can get the simple bound
$$\frac{1+\oo(1)}{[1 - (x\rho)^{-1}]^3} \sum_{i=1}^2 \bigbrace{\sqrt{\frac{\rho_i}{\rho}} + \sqrt{\frac{1}{x\rho}}}^4 < 5. $$
Inserting it into \eqref{solval1}, we can solve for the upper bound in \eqref{eq_uplowx}.
%where we abbreviated $\rho:=\rho_1+\rho_2$.

We can get better bounds if the values of $\rho_1$ and $\rho_2$ increase. For example, if we consider the case $\min(\rho_1,\rho_2)\ge 100$, then with some basic calculations, one can show that in this case
$$ \frac{1}{[1 - (x \rho)^{-1}]^3} \sum_{i=1}^2 \bigbrace{\sqrt{\frac{\rho_i}{\rho}} + \sqrt{\frac{1}{x \rho}}}^4 <  \frac{\rho_1^2 + \rho_2^2}{\rho^2} + 0.52.$$
Thus the following inequality implies \eqref{solval1}:
\be\nonumber %\label{solval1add}
\begin{split}
&\left(  \frac{\rho_1^2 + \rho_2^2}{\rho^2} + 0.52\right) \Psi(\beta_1,\beta_2) < \frac{1}{\rho_1 -1}+\frac{1}{\rho_2 -1} - \frac{2 }{x\rho - 1} ,
\end{split}
\ee
%In particular, if 
%\be\nonumber
%\begin{split}
%&\left(  \frac{\rho_1^2 + \rho_2^2}{(\rho_1+\rho_2)^2} + 0.32\right) pd^2 < \frac{\sigma^2}{\rho_1 -1}+\frac{\sigma^2}{\rho_2 -1} - \frac{2\sigma^2}{(\rho_1+\rho_2) - 1} ,
%\end{split}
%\ee
%that is, we have positive transfer when using all the data, then 
from which we can solve for the following upper bound on $x^\star$:
\begin{align*}
x^\star &<  \frac1{\rho} \frac{2 }{\frac{1}{\rho_1-1}+\frac1{\rho_2-1}  -  \left(  \frac{\rho_1^2 + \rho_2^2}{\rho^2} + 0.52\right)\Psi(\beta_1,\beta_2)}+ \frac1\rho .
%\\
%& < \frac1{\rho_1+\rho_2}\left[\frac{2 }{\frac{1}{\rho_1}+\frac1{\rho_2}  - \left(  \frac{\rho_1^2 + \rho_2^2}{(\rho_1+\rho_2)^2} + \frac13\right)\frac{pd^2}{\sigma^2}}+ 1\right] .
\end{align*}
Similarly, we can get a better lower bound. 
%\noindent{\bf The lower bound.} 
From \eqref{reduceproof1} and \eqref{reduceproof2}, we see that $x^\star\ge x$ if $x$ satisfies
\be\label{solval2}
\begin{split} 
&(1-\oo(1)) \cdot  \frac{\Psi(\beta_1,\beta_2)}{[1 - (x \rho)^{-1}]^3} \sum_{i=1}^2 \bigbrace{\sqrt{\frac{\rho_i}{\rho}} - \sqrt{\frac{1}{x \rho}}}^4 \ge \frac{1}{\rho_1 -1}+\frac{1}{\rho_2 -1} - \frac{2 }{x \rho - 1} . 
\end{split}
\ee
%We then follow similar arguments as the above proof for the upper bound.
Then in the case $\min(\rho_1,\rho_2)\ge 100$, with some basic calculations, one can show that the sum on the left-hand side of \eqref{solval2} satisfies 
$$ \frac{1}{[1 - (x \rho)^{-1}]^3} \sum_{i=1}^2 \bigbrace{\sqrt{\frac{\rho_i}{\rho}} - \sqrt{\frac{1}{x\rho}}}^4 >  \frac{\rho_1^2 + \rho_2^2}{\rho^2} -0.33 .$$
Thus the following inequality implies \eqref{solval2}:
\be\label{solval2add}
\begin{split}
&\left( \frac{\rho_1^2 + \rho_2^2}{\rho^2} -0.33\right) pd^2 > \frac{\sigma^2}{\rho_1 -1}+\frac{\sigma^2}{\rho_2 -1} - \frac{2\sigma^2}{x\rho - 1} ,
\end{split}
\ee
%There are two cases: if 
%\be\nonumber
%\begin{split}
%&\left( \frac{\rho_1^2 + \rho_2^2}{(\rho_1+\rho_2)^2} -0.26\right) pd^2 \ge \frac{\sigma^2}{\rho_1 -1}+\frac{\sigma^2}{\rho_2 -1},
%\end{split}
%\ee
%then we have negative transfer for all choice of $0\le \al \le 1$; otherwise, 
from which we can solve for the following lower bound on $x^\star$:
\begin{align*}
x^\star &>  \frac1\rho \frac{2 }{\frac{1}{\rho_1-1}+\frac1{\rho_2-1}  - \left( \frac{\rho_1^2 + \rho_2^2}{\rho^2} -0.33\right) \Psi(\beta_1,\beta_2)}+ \frac1\rho .
\end{align*}
This gives a lower bound above $x_l$.
\end{proof}

@@

We now prove Proposition \ref{prop_covariate}, which shows that $\te(\hat{\beta}^{\MTL})$ is minimized approximately when $M$ is a scalar matrix where there is enough source data.

\begin{proof}[Proof of Proposition \ref{prop_covariate}]
Let 
$$M_0:=\argmin_{M\in \cal S_{\mu}}g(M).$$ 
We now calculate $g(M_0)$. With the same arguments as in Lemma \ref{lem_hat_v}, we can show that \eqref{hatv_add1} holds. Moreover, if the parameters are chosen such that $p^{-1+c_0} \sigma^2 \le \kappa^2  \le p^{-\e_0-c_0}\sigma^2$ as in \eqref{choiceofpara}, we can simplify
\be \nonumber
\begin{split}
g(M_0)&=(1+\OO(p^{-\e}))\cdot \sigma^2  \bigtr{\Sigma_2(X_1^{\top}X_1  + X_2^{\top}X_2)^{-1} }  ,
\end{split}
\ee
with high probability for some constant $\e>0$. In fact, Lemma \ref{lem_hat_v} was proved assuming that $M=\id$, but its proof can be easily extended to the case with general $M\in \cal S_{\mu}$ by using that $ \mu_{\min}\le \lambda_p(M)\le \lambda_1(M)\le \mu_{\max}$. We omit the details here.

>>>>

\begin{proof}[Proof of Proposition \ref{prop_data_size}]
Following the proof of Proposition \ref{prop_dist_transition}, we know that $\te(\hat{\beta}_2^{\MTL})< \te(\hat{\beta}_2^{\STL})$ w.h.p. if \eqref{pos1} holds, while $\te(\hat{\beta}_2^{\MTL})>\te(\hat{\beta}_2^{\STL})$ w.h.p. if \eqref{neg1} holds. 
%\be\label{pos1}
%pd^2  \cdot \frac{\rho_1 (\rho_1 + \rho_2) (\rho_2 - 1)}{(\rho_1 + \rho_2 - 1)^2}\bigbrace{1 + \sqrt{\frac{1}{\rho_1}}}^4 <  \sigma^2  \bigbrace{1 - \oo(1)},\ee
%and we get from \eqref{upper102} that $\te(\hat{\beta}_t^{\MTL})>\te(\hat{\beta}_t^{\STL})$ w.h.p. if
%\be\label{neg1}pd^2  \cdot \frac{\rho_1 (\rho_1 + \rho_2) (\rho_2 - 1)}{(\rho_1 + \rho_2 - 1)^2}\bigbrace{1 - \sqrt{\frac{1}{\rho_1}}}^4 >  \sigma^2  \bigbrace{1 + \oo(1)}.\ee
%then we have $\te(\hat{\beta}_t^{\MTL})> \te(\hat{\beta}_t^{\STL})$ w.h.p..

We first explain the meaning of the condition 
\be\label{explainpsi}\Psi(\beta_1,\beta_2)>2/(\rho_2-1).\ee 
%For the first statement of Proposition \ref{prop_data_size}, we 
Notice that the function
$$ \Phi(\rho_1, \rho_2)=\frac{(\rho_1 + \rho_2 - 1)^2}{\rho_1 (\rho_1 + \rho_2) (\rho_2 - 1)}=\frac{1}{\rho_2-1} \left(1 +\frac{\rho_2-2}{\rho_1}+\frac{1}{\rho_1(\rho_1+\rho_2)}\right)$$
is strictly decreasing with respect to $\rho_1$ as long as $\rho_2> 2$, and $ \Phi(\rho_1, \rho_2)$ converges to $(\rho_2-1)^{-1}$ as $\rho_1\to \infty$. Since $\left( \al_-^2(\rho_1) -  \oo(1) \right)^{-1} < 2$ for $\rho_1>40$,  \eqref{explainpsi} implies that \eqref{neg1} holds for large enough $\rho_1$. Moreover, the transition from positive transfer when $\rho_1$ is small to negative transfer when $\rho_1$ is large is described by the two bounds in Proposition \ref{prop_data_size}, which we will show now. 
% $$ \Psi(\beta_1,\beta_2) \ge \bigbrace{1 + \sqrt{\frac{1}{\rho_1}}}^{-4}\frac{\sigma^2}{\rho_2-1}\left(1 -\oo(1)\right),$$ 
% then \eqref{pos1} holds, which shows that  $\te(\hat{\beta}_t^{\MTL}) < \te(\hat{\beta}_t^{\STL})$ holds w.h.p.. Plugging into $\rho_1>50$, we obtain the first statement. ...... Note that for $\rho_1>40$, we have 
%$\gamma_+^{-1}$
%The two bounds follows directly from \eqref{pos1} and \eqref{neg1}. 
In the following proof, we will use the following trivial inequalities 
\be\label{trivialphi}
\frac{(\rho_2-1) \rho_1 }{\rho_1+\rho_2-2}\cdot \left( 1-\frac{1}{(\rho_1+\rho_2-2)^2}\right) \le  \frac1{\Phi(\rho_1, \rho_2)} \le  \frac{(\rho_2-1) \rho_1 }{\rho_1+\rho_2-2}.
\ee

@@

 Note that if we require the RHS of \eqref{addconstraint} to be larger than $40$, that is, \eqref{addconstraint} is not a null condition, then with \eqref{explainpsi} we get
 $$ \rho_2 - 2>\left[ 2\left( \al_+^2(\rho_1) +  \oo(1) \right)  -1\right]\rho_1 .$$
Plugging into $\rho_1>40$, we get $\rho_2 \ge 106$. This gives a constraint on $\rho_2$.

@@

 \noindent{\bf Negative transfer.} 
With \eqref{trivialphi}, we see that \eqref{neg1} is implied by the following inequality:
\begin{align}\label{neg1solv}
 & \Psi(\beta_1,\beta_2)  \frac{(\rho_2-1) \rho_1 }{\rho_1+\rho_2-2}\left( 1-\frac{1}{(\rho_1+\rho_2-2)^2}\right) >  \Psi(\beta_1,\beta_2) \frac{(\rho_2-1.5) \rho_1 }{\rho_1+\rho_2-2}> \left( \al_-^2(\rho_1) -  \oo(1) \right)^{-1} .
 \end{align}
 where we used $(1-(\rho_1+\rho_2-2)^{-2})(\rho_2 - 1) > \rho_2-1.5$ for $\rho_1>40$ and $\rho_2>110$. Then we can solve \eqref{neg1solv} to get that
 \be\label{addconstaint2}
 \rho_1 > \frac{\rho_2-2 }{\Psi(\beta_1, \beta_2) \cdot  (\rho_2 - 1.5)\left( \al_-^2(\rho_1) -  \oo(1) \right)  - 1}  ,
 \ee
which gives the second statement of Proposition \ref{prop_data_size}.
We remark that condition \eqref{explainpsi} implies $\Psi(\beta_1, \beta_2) \cdot  (\rho_2 - 1.5)\left( \al_-^2(\rho_1) -  \oo(1) \right)>1$, so \eqref{addconstaint2} does not give a trivial bound. 
 \end{proof}

Next we prove Proposition \ref{prop_data_efficiency}, which gives precise upper and lower bounds on the data efficiency ratio for Taskonomy.

\begin{proof}[Proof of Proposition \ref{prop_data_efficiency}]
Suppose we have reduced number of datapoints---$x n_1$ for task 1 and $x n_2$ for task 2 with $n_1=\rho_1 p$ and $n_2=\rho_2 p$. Then all the results in the proof of Proposition \ref{prop_dist_transition} still hold, except that we need to replace $(\rho_1,\rho_2)$ with $(x\rho_1,x\rho_2)$. More precisely, we have
	\begin{align*}
		 a_1 = \frac{\rho_1(x\rho_1 + x\rho_2 - 1)}{x(\rho_1 + \rho_2)^2} ,\quad
		& a_2 = \frac{\rho_2(x\rho_1 + x\rho_2 - 1)}{x(\rho_1 + \rho_2)^2} ,  \\
		 a_3 = \frac{\rho_2}{(\rho_1 + \rho_2)(x\rho_1 + x\rho_2 - 1)}, \quad
		& a_4 = \frac{\rho_1}{(\rho_1 + \rho_2)(x\rho_1 + x\rho_2 - 1)}. 
	\end{align*}
Then with Lemma \ref{lem_cov_shift} we can obtain that with high probability,
	\be\label{reduceproof1}
\begin{split}
\te_i(\hat \beta_i^{\MTL}(x)) =  \frac{\sigma^2}{x (\rho_1+\rho_2) - 1}\left(1+ \oo(1)\right)+ \delta^{(i)}_{\bias}, \quad i=1,2,
\end{split}
\ee
 where as in \eqref{Deltabeta}, the model shift biases $\delta^{(i)}_{\bias}$ satisfy that with high probability,
\be\nonumber %\label{Deltabeta} 
\al_-^2(\al\rho_i) - \oo(1)  \le {\delta_{\bias}^{(i)}}/{ \Delta_{\bias}^{(i)}} \le \al_+^2(\al\rho_i) +  \oo(1) , \quad i=1,2.\ee	
Here $ \Delta^{(i)}_{\bias}$ are defined as 
\be \nonumber
\Delta^{(i)}_{\bias} := pd^2 \frac{(x\rho_i)^2\cdot x (\rho_1+\rho_2)}{[x ( \rho_1+\rho_2) - 1]^3} ,\quad i=1,2,. 
\ee
On the other hand, by Lemma \ref{lem_minv} we have that w.h.p.,
\be\label{reduceproof2} 
\te_i(\hat{\beta}_i^{\STL}) = \frac{\sigma^2}{\rho_i -1} \left( 1+ \oo(1)\right),\quad i=1,2.
\ee
Comparing \eqref{reduceproof1} and \eqref{reduceproof2}, we immediately obtain the lower bound $x^\star\ge x_l $. 
%where
%$$\al_l:=\frac1{\rho_1+\rho_2}\left(\frac{2}{\frac1{\rho_1-1}+\frac1{\rho_2 -1}}+1\right) \ge \frac{\min(\rho_1,\rho_2)}{\rho_1+\rho_2}.$$
In fact, if $x< x_l$, then we have that
$$ \frac{2\sigma^2}{x (\rho_1+\rho_2) - 1} > \frac{\sigma^2}{\rho_1-1}+\frac{\sigma^2}{\rho_2-1}.$$
This shows that $\te_1(\hat{\beta}_1^{\MTL}(x)) + \te_2(\hat{\beta}_2^{\MTL}(x))$ is larger than $\te_1(\hat{\beta}_1^{\STL}) + \te_2(\hat{\beta}_2^{\STL})$ even if we do not take into account the model shift biases $ \delta^{(i)}_{\bias}$.

@@

%\noindent{\bf The upper bound.} 
From \eqref{reduceproof1} and \eqref{reduceproof2}, we see that $x^\star\le x$ if $x$ satisfies 
\be\nonumber
\begin{split}
&(1+\oo(1)) \cdot \sum_{i=1}^2 pd^2 \frac{(x\rho_i)^2\cdot x\rho }{(x\rho - 1)^3}  \bigbrace{1 + \sqrt{\frac{1}{x \rho_i}}}^4 \le \frac{\sigma^2}{\rho_1 -1}+\frac{\sigma^2}{\rho_2 -1} - \frac{2\sigma^2}{x\rho  - 1} . 
\end{split}
\ee
We rewrite the inequality as
\be
\begin{split}\label{solval1}
  (1+\oo(1)) \cdot \frac{\Psi(\beta_1,\beta_2)}{[1 - (x\rho)^{-1}]^3} \sum_{i=1}^2 \bigbrace{\sqrt{\frac{\rho_i}{\rho}} + \sqrt{\frac{1}{x\rho}}}^4\le \frac{1}{\rho_1 -1}+\frac{1}{\rho_2 -1} - \frac{2}{x \rho  - 1}.
\end{split}
\ee
With $x\rho\ge \min(\rho_1,\rho_2)\ge 9$, we can get the simple bound
$$\frac{1+\oo(1)}{[1 - (x\rho)^{-1}]^3} \sum_{i=1}^2 \bigbrace{\sqrt{\frac{\rho_i}{\rho}} + \sqrt{\frac{1}{x\rho}}}^4 < 5. $$
Inserting it into \eqref{solval1}, we can solve for the upper bound in \eqref{eq_uplowx}.
%where we abbreviated $\rho:=\rho_1+\rho_2$. 
\end{proof}
\begin{remark}
We can get better bounds if the values of $\rho_1$ and $\rho_2$ increase. For example, if we consider the case $\min(\rho_1,\rho_2)\ge 100$, then with some basic calculations, one can check that in this case
$$ \frac{1}{[1 - (x \rho)^{-1}]^3} \sum_{i=1}^2 \bigbrace{\sqrt{\frac{\rho_i}{\rho}} + \sqrt{\frac{1}{x \rho}}}^4 <  \frac{\rho_1^2 + \rho_2^2}{\rho^2} + 0.52.$$
Thus the following inequality implies \eqref{solval1}:
\be\nonumber %\label{solval1add}
\begin{split}
&\left(  \frac{\rho_1^2 + \rho_2^2}{\rho^2} + 0.52\right) \Psi(\beta_1,\beta_2) < \frac{1}{\rho_1 -1}+\frac{1}{\rho_2 -1} - \frac{2 }{x\rho - 1} ,
\end{split}
\ee
%In particular, if 
%\be\nonumber
%\begin{split}
%&\left(  \frac{\rho_1^2 + \rho_2^2}{(\rho_1+\rho_2)^2} + 0.32\right) pd^2 < \frac{\sigma^2}{\rho_1 -1}+\frac{\sigma^2}{\rho_2 -1} - \frac{2\sigma^2}{(\rho_1+\rho_2) - 1} ,
%\end{split}
%\ee
%that is, we have positive transfer when using all the data, then 
from which we can solve for the following upper bound on $x^\star$:
\begin{align*}
x^\star &<  \frac1{\rho} \frac{2 }{\frac{1}{\rho_1-1}+\frac1{\rho_2-1}  -  \left(  \frac{\rho_1^2 + \rho_2^2}{\rho^2} + 0.52\right)\Psi(\beta_1,\beta_2)}+ \frac1\rho .
%\\
%& < \frac1{\rho_1+\rho_2}\left[\frac{2 }{\frac{1}{\rho_1}+\frac1{\rho_2}  - \left(  \frac{\rho_1^2 + \rho_2^2}{(\rho_1+\rho_2)^2} + \frac13\right)\frac{pd^2}{\sigma^2}}+ 1\right] .
\end{align*}
Similarly, we can get a better lower bound. 
%\noindent{\bf The lower bound.} 
From \eqref{reduceproof1} and \eqref{reduceproof2}, we see that $x^\star\ge x$ if $x$ satisfies
\be\label{solval2}
\begin{split} 
&(1-\oo(1)) \cdot  \frac{\Psi(\beta_1,\beta_2)}{[1 - (x \rho)^{-1}]^3} \sum_{i=1}^2 \bigbrace{\sqrt{\frac{\rho_i}{\rho}} - \sqrt{\frac{1}{x \rho}}}^4 \ge \frac{1}{\rho_1 -1}+\frac{1}{\rho_2 -1} - \frac{2 }{x \rho - 1} . 
\end{split}
\ee
%We then follow similar arguments as the above proof for the upper bound.
Then in the case $\min(\rho_1,\rho_2)\ge 100$, with some basic calculations, one can show that the sum on the left-hand side of \eqref{solval2} satisfies 
$$ \frac{1}{[1 - (x \rho)^{-1}]^3} \sum_{i=1}^2 \bigbrace{\sqrt{\frac{\rho_i}{\rho}} - \sqrt{\frac{1}{x\rho}}}^4 >  \frac{\rho_1^2 + \rho_2^2}{\rho^2} -0.33 .$$
Thus the following inequality implies \eqref{solval2}:
\be\label{solval2add}
\begin{split}
&\left( \frac{\rho_1^2 + \rho_2^2}{\rho^2} -0.33\right) \Psi(\beta_1,\beta_2)  > \frac{1}{\rho_1 -1}+\frac{1}{\rho_2 -1} - \frac{2 }{x\rho - 1} ,
\end{split}
\ee
%There are two cases: if 
%\be\nonumber
%\begin{split}
%&\left( \frac{\rho_1^2 + \rho_2^2}{(\rho_1+\rho_2)^2} -0.26\right) pd^2 \ge \frac{\sigma^2}{\rho_1 -1}+\frac{\sigma^2}{\rho_2 -1},
%\end{split}
%\ee
%then we have negative transfer for all choice of $0\le \al \le 1$; otherwise, 
from which we can solve for the following lower bound on $x^\star$:
\begin{align*}
x^\star &>  \frac1\rho \frac{2 }{\frac{1}{\rho_1-1}+\frac1{\rho_2-1}  - \left( \frac{\rho_1^2 + \rho_2^2}{\rho^2} -0.33\right) \Psi(\beta_1,\beta_2)}+ \frac1\rho .
\end{align*}
This gives a lower bound above $x_l$.
\end{remark}

@@

Finally, in this subsection we prove Proposition \ref{prop_covariate}, which shows that $\te(\hat{\beta}_2^{\MTL})$ is minimized approximately when $M$ is a scalar matrix, provided that there is enough source data.

\begin{proof}[Proof of Proposition \ref{prop_covariate}]
Denote the minimizer of $g$ by 
$$M_0:=\argmin_{M\in \cal S_{\mu}}g(M).$$ 
We now calculate $g(M_0)$. With the same arguments as in Claim \ref{lem_hat_v}, we can show that \eqref{hatv_add1} holds. Moreover, if the parameters are chosen such that $p^{-1+c_0} \sigma^2 \le \kappa^2  \le p^{-\e_0-c_0}\sigma^2$ as in \eqref{choiceofpara}, we can simplify $g(M_0)$ as
\be \nonumber
\begin{split}
g(M_0)&=(1+\OO(p^{-\e}))\cdot \sigma^2  \bigtr{\Sigma_2(X_1^{\top}X_1  + X_2^{\top}X_2)^{-1} }  \quad \text{w.h.p.}
\end{split}
\ee
 for some constant $\e>0$. In fact, Lemma \ref{lem_hat_v} was proved assuming that $M=\id$, but its proof can be easily extended to the case with general $M\in \cal S_{\mu}$ by using that $ \mu_{\min}\le \lambda_p(M)\le \lambda_1(M)\le \mu_{\max}$. We omit the details here.
